### 2015-03-12 beamer_test.tex
added beamer testfile

\documentclass{beamer}
\usetheme{Frankfurt}
\usepackage[ngerman]{babel}
\usepackage{PDBF}
%\setbeameroption{show only notes}

\mdqoff

\addtobeamertemplate{note page}{\setlength{\parskip}{12pt}}
\makeatletter
\beamertemplatenavigationsymbolsempty
\setbeamerfont{footline}{size=\normalsize}
\setbeamertemplate{footline}{{\par\hspace*{\fill}\vspace*{0.25cm}\insertframenumber~~~}}
\setbeamertemplate{items}[default]
\setbeamertemplate{sections/subsections in toc}[sections numbered]
\setbeamertemplate{subsection in toc}{}
\setbeamercolor{title}{fg=white,bg=red!50!green!70}
\setbeamercolor{frametitle}{fg=white,bg=red!50!green!70}
\setbeamercolor{subsection in toc}{fg=black}
\setbeamercolor{section in toc}{fg=black}
\setbeamercolor{item}{fg=black}
\makeatother

\title{Beamer and PDBF}   
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
  \\[5pt]
 Christian Goldstein\\
  \texttt{c.goldstein@goldstein-systemberatung.de}
}
\date{}

\DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
\DBSQLFile{somedb.sql}
\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment

\begin{document}

\begin{frame}
\titlepage
\vspace{-2.3cm}
\begin{center}
\includegraphics[scale=0.25]{C:/Users/User/Downloads/UdS-Logo.jpg}
\end{center}
Presenter: Patrick Bender
\end{frame}
\note{bla}

\section{SQL Queries} 
\subsection*{}
\begin{frame}\frametitle{Simple Query}

\end{frame}
\note{bla}

\subsection*{}
\begin{frame}\frametitle{Advanced Query}

\end{frame}
\note{bla}

\section{Charts} 
\subsection*{}
\begin{frame}\frametitle{Line Chart with little data}
\def\test{testtest}
	{\centering
	\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=\test Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
	\\}
\end{frame}
\note{bla}

\subsection*{}
\begin{frame}\frametitle{Line Chart with much data}
	{\centering
	\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
	\\}
\end{frame}
\note{bla}

\section*{}
\begin{frame}
	Thank you for your attention!
\end{frame}
\note{bla}

\begin{frame}
	Questions?
\end{frame}
\note{bla}

\end{document}