### 2015-11-30 src/pdbf/tests/charts.tex

\documentclass[10pt]{article}
\usepackage{pdbf}
\usepackage[a4paper]{geometry}

\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 10, 15); INSERT INTO test VALUES(2, 100, 90); INSERT INTO test VALUES(3, 120, 110); INSERT INTO test VALUES(4, 100, 150);}

\def\test{
	%
	%%%%%%%%%%%%%%%%%%% Page1
	%
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"tick": { "rotate": 45 }
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"tick": { "rotate": 90 }
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-left",
				"x": 25,
				"y": 0,
				"step": 1
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-left",
				"x": 25,
				"y": 0
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-right",
				"step": 1
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-right"
			}
		}
	}]{SELECT * FROM test;}\\
	%
	%%%%%%%%%%%%%%%%%%% Page2
	%
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "right"
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"show": false
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "inner-center"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "inner-middle"
				}
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "inner-left"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "inner-bottom"
				}
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "outer-right"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "outer-top"
				}
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "outer-center"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "outer-middle"
				}
			}
		}
	}]{SELECT * FROM test;}\\
}

\begin{document}
\noindent
\tiny
\test
\normalsize
\test
\LARGE
\test
%
\fontfamily{phv}
\tiny
\test
\normalsize
\test
\LARGE
\test
\end{document}